Characterizing Stellar Streams by ML

### 2023-08-29 src/tex/ms.tex
[showyourwork] overleaf sync

    \begin{figure*}
        \script{mock/plot/results.py}
        \centering
        \includegraphics[width=1\linewidth]{figures/mock/results.pdf}
        \caption{
            %
            \textbf{Top panel:}%
                The stream mixture coefficient $f(\phi_1)$ (black dotted line)
                predicted by the model, overplotted on the true stream fraction
                (black-outlined histogram), calculated as the ratio of stream
                stars to total stars in 55 bins over the $\phi_1$ range of the
                isochrone.  \textbf{Middle panel: }%
                $\phi_2(\phi_1)$ over the full range of $\phi_1$. The data color
                and transparency is set by the model's MLE stream probability
                (with dropout turned off). Plotted behind the data are the true
                labels for the stream stars (large black circles).  The true
                labels exactly limn the model-identified stream stars.  The MLE
                stream track (black dashed line) is over-plotted on the
                probability-colored data.
            %
            \textbf{Lower panel 3: }%
                $\varpi_2(\phi_1)$ over the full range of $\phi_1$. Like for
                $\phi_2$, the data color and transparency is set by the MLE
                probability, with the true labels (large black circles)
                under-plotted and MLE track (black dashed line) over-plotted.
            %
            \textbf{Upper Histograms: }%
                $\phi_2$ histograms for each model component (blue background,
                red stream), in 4 $\phi_1$ ranges across the data set.  The
                height of each bin is the probability-weighted density.  The
                background is a uniform distribution, while the stream is a
                Gaussian. In projection across the $\phi_1$ range it can appear
                multi-modal (panel 2) if there are large gaps or `smeared' out
                if the Gaussian's mean changes significantly over the range.
            %
            \textbf{Upper Middle Histograms: }%
                Parallax $\varpi$ histograms for each model component (blue
                background, red stream), in 4 $\phi_1$ ranges across the data
                set.  The height of each bin is the probability-weighted
                density.
            %
            \textbf{Lower Middle and Lower Histograms:}%
                Photometric coordinate $(g, r)$ plots for each model component
                in 4 $\phi_1$ ranges across the data set. The top row plots the
                stream stars (red) over the background. The lower row plots the
                background stars (blue) over the stream stars to show that the
                how the two distributions mix.  } TODO:
        \begin{enumerate}
            \item the linear density in the stream-fraction plot
        \end{enumerate}
        \label{fig:mock_data_result}
    \end{figure*}

@@

    We find that the stream is successfully recovered in both astrometric and
    photometric coordinates, with $\variable{output/mock/ncorrect_variable.txt}\%$ stream
    stars being recovered with membership probability greater than $80\%$. The false
    positive rate (i.e., stars misidentified as stream members) for this test is
    found to be $\variable{output/mock/falseident_variable.txt}\%$ when cutting on
    this threshold.

    In the bottom two rows of \autoref{fig:mock_data_result} we illustrate the
    synthetic data color-coded by stream membership probability (top) and
    background membership probability (background) for different $\phi_1$
    slices. Importantly, the photometric portion of the mixture model
    successfully identifies the stream in magnitude space, showing a high stream
    membership probability where the stream is present, and a low probability
    ($\sim 0$) where the stream is absent.

@@

    We now apply our density modeling approach to the stellar stream GD-1. We
    discuss the data selection in \autoref{sub:results_gd1_data} and results in
    \autoref{ssub:trained_gd1}.

    \subsection{Data Selection} \label{sub:results_gd1_data}

@@

    \subsection{Model Specification}\label{sec:GD_1_ModelSpecification}

@@

    \subsection{Trained Density Model}\label{ssub:trained_gd1}

>>>>

    \begin{figure*}
        \script{mock/plot/results.py}
        \centering
        \includegraphics[width=1\linewidth]{figures/mock/results.pdf}
        \caption{
            %
            \textbf{Top panel:}%
                The stream mixture coefficient $f(\phi_1)$ (black dotted line)
                predicted by the model, overplotted on the true stream fraction
                (black-outlined histogram), calculated as the ratio of stream
                stars to total stars in 55 bins over the $\phi_1$ range of the
                isochrone.  \textbf{Middle panel: }%
                $\phi_2(\phi_1)$ over the full range of $\phi_1$. The data color
                and transparency is set by the model's MLE stream probability
                (with dropout turned off). Plotted behind the data are the true
                labels for the stream stars (large black circles).  The true
                labels exactly limn the model-identified stream stars.  The MLE
                stream track (black dashed line) is over-plotted on the
                probability-colored data.
            %
            \textbf{Lower panel 3: }%
                $\varpi_2(\phi_1)$ over the full range of $\phi_1$. Like for
                $\phi_2$, the data color and transparency is set by the MLE
                probability, with the true labels (large black circles)
                under-plotted and MLE track (black dashed line) over-plotted.
            %
            \textbf{Upper Histograms: }%
                $\phi_2$ histograms for each model component (blue background,
                red stream), in 4 $\phi_1$ ranges across the data set.  The
                height of each bin is the probability-weighted density.  The
                background is a uniform distribution, while the stream is a
                Gaussian. In projection across the $\phi_1$ range it can appear
                multi-modal (panel 2) if there are large gaps or `smeared' out
                (panel 3) if the Gaussian's mean changes significantly over the range.
            %
            \textbf{Upper Middle Histograms: }%
                Parallax $\varpi$ histograms for each model component (blue
                background, red stream), in 4 $\phi_1$ ranges across the data
                set.  The height of each bin is the probability-weighted
                density.
            %
            \textbf{Lower Middle and Lower Histograms:}%
                Photometric coordinate $(g, r)$ plots for each model component
                in 4 $\phi_1$ ranges across the data set. The top row plots the
                stream stars (red) over the background. The lower row plots the
                background stars (blue) over the stream stars to show that the
                how the two distributions mix.
            % 
        }
        \label{fig:mock_data_result}
        TODO:
        \begin{enumerate}
            \item the linear density in the stream-fraction plot
        \end{enumerate}
    \end{figure*}

@@

    We find that the stream is successfully recovered in both astrometric and
    photometric coordinates, with $\variable{output/mock/ncorrect_variable.txt}\%$ stream
    stars being recovered with membership probability greater than $80\%$.
    In the bottom two rows of \autoref{fig:mock_data_result} we illustrate the
    synthetic data color-coded by stream membership probability (top) and
    background membership probability (background) for different $\phi_1$
    slices. Importantly, the photometric portion of the mixture model
    successfully identifies the stream in magnitude space, showing a high stream
    membership probability where the stream is present, and a low probability
    ($\sim 0$) where the stream is absent.
    The false positive rate (i.e., stars misidentified as stream members) for this test
    is found to be $\variable{output/mock/falseident_variable.txt}\%$ when cutting on
    the 80\% membership probability threshold.
    % The perfect recovery rate and low false identification rate demonstrate the efficacy
    % of this method.

    % Discussion why not finding the small-scale fluctuations?

@@

    Having successful applied the modeling framework to a mock stream,
    we now apply our density modeling approach to the stellar stream GD-1.
    We
    discuss the data selection in \autoref{sub:results_gd1:data},
    model specifics in \autoref{sub:results_gd1:model},
    and the results in \autoref{sub:results_gd1:results}.

    \subsection{Data Selection} \label{sub:results_gd1:data}

@@

    \subsection{Model Specification}\label{sub:results_gd1:model}

@@

    \subsection{Trained Density Model}\label{sub:results_gd1:results}
